Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-09-17 paper/abstract.tex
start on revising first part of paper

Cyber-physical systems, such as avionics, must be tolerant to
cyber-attacks in the same way they are tolerant to random faults: gracefully recover or safely shut down as requirements dictate.
\brfcs\ is a model-based systems engineering environment that extends the Open Source AADL Tool Environment to add new design, analysis, and code generation capabilities for building cyber-resilient systems.
This manuscript describes a mechanism implemented by \brfcs\ to create, and insert, high-assurance components into a system for cyber-resiliency.
Such components can be filters that guard against malformed input or monitors that guard against spoofing and other malicious behavior.
A formal specification in the form of a code contract defines each high-assurance component and is used with test contracts to verify the specified behavior.
Once tested, the code contract is used to verify that the added component addresses system-level cyber requirements.
Implementations for these
high-assurance components are directly synthesized from their code
contracts and are backed up by proofs showing that high-level
specifications map in a meaning-preserving way to code.
The manuscript further reports on a case study that used \brfcs\ to cyber-harden a UAV system by inserting high assurance components to isolate the open source Air Force Research Laboratory's OpenUxAS services for
route planning.  The case study
demonstrates that synthesizing implementations from code contracts is feasible in real-world systems
engineering.

>>>>

Cyber-physical systems, such as avionics, must be tolerant to
cyber-attacks in the same way they are tolerant to random faults: they
must gracefully recover, or safely shut down, as requirements dictate.
We have developed a workflow for creating, and inserting,
high-assurance components implementing cyber-resiliency into a
model-based systems engineering environment.  Example high-assurance
components are filters, which guard against malformed input, and
runtime monitors, which guard against spoofing and other malicious
behavior. A formal specification in the form of a \emph{code contract}
defines each high-assurance component and is developed with the
support of \emph{test contracts} for testing the specified behavior.
Once tested, model checking used to verify that the added
high-assurance component indeed addresses system-level cyber
requirements.  Implementations for these high-assurance components are
directly synthesized from their code contracts and are backed up by
proofs showing that high-level specifications map in a
semantics-preserving way to code generated by a verified compiler.  We
report on a case study that cyber-hardened a UAV system by inserting
high assurance components to harden the open source Air Force Research
Laboratory's OpenUxAS services for route planning.  The case study
demonstrates that synthesizing correct implementations from code
contracts is feasible in real-world systems engineering.

### 2022-09-17 paper/intro.tex
start on revising first part of paper

\egm{Add to introduction and related work, that what is done at Galois called Copilot. Its a stream processing langague that generates code from the specification. Copilot to C. Effectively CodeGen from Lustre. Bounds how far needed to look in past for any value.}
In recent years, aerospace stakeholders have realized that avionics systems are subject to possible cyber-attacks just like other cyber-physical systems.  
Thus, in addition to being fault-tolerant, safety-critical avionics systems must also be {\em cyber-resilient}.
Cyber-resiliency means that the system is tolerant to cyber-attacks just as safety-critical systems are tolerant to random faults: they recover and continue to execute their mission function, or safely shut down, as requirements dictate.

@@

\egm{Remove DARPA junk.}
The DARPA Cyber Assured Systems Engineering (CASE) project is targeted at developing tools for design, analysis, and verification that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{This work was funded in part by the Defense Advanced Research Projects Agency (DARPA).  The views expressed are those of the authors and do not reflect the official policy or position of DARPA or the U.S. Government.}
One such tool is the Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities for building cyber-resilient systems.

This manuscript details how \brfcs\ facilitates inserting, specifying, testing, and synthesizing high assurance components into a system to improve its cyber-resiliency.
These \egm{make clear what these refers to} are referred to as \emph{transforms} by \brfcs.
Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components.
\brfcs\ automates inserting the components into the system, and the designer writes a \emph{code contract} for each component to define its intended behavior.

@@

Another novel aspect of the approach discussed in this manuscript is the synthesis of the code contracts for the high-assurance components to \ckml, a verified
compiler implementation for the functional programming language
ML \cite{cakeml}. 
This manuscript describes in detail the synthesis path
from code contracts to \ckml\ code, providing a formal framework in
which to argue correctness. \ckml\ then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the \ckml\ source code is exactly
preserved in the final binaries). 
A key contribution here is that the code contract semantics are defined in such a way that the \agr\ verification 
results for the code contract hold for the deployed component, \ie, the component will detect
and prevent the indicated cyber-vulnerabilities over all possible
finite inputs. Preliminary work has shown how to lift this result to
infinite input traces as these systems are inherently reactive and
intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

>>>>

\egm{Add to introduction and related work, that what is done at
Galois called Copilot. Its a stream processing langague that generates
code from the specification. Copilot to C. Effectively CodeGen from
Lustre. Bounds how far needed to look in past for any value.}

In recent years, aerospace stakeholders have realized that avionics
systems are subject to possible cyber-attacks just like other
cyber-physical systems.
%
\footnote{This work was funded in part by the
Defense Advanced Research Projects Agency (DARPA).  The views
expressed are those of the authors and do not reflect the official
policy or position of DARPA or the U.S. Government.}
%
Thus, in addition to being fault-tolerant, safety-critical avionics
systems must also be {\em cyber-resilient}.  Cyber-resiliency means
that the system is tolerant to cyber-attacks just as safety-critical
systems are tolerant to random faults: they recover and continue to
execute their mission function, or safely shut down, as requirements
dictate.

@@

%% The DARPA Cyber Assured Systems Engineering (CASE) project is targeted
%% at developing tools for design, analysis, and verification that enable
%% systems engineers to {\em design-in} cyber-resiliency for complex
%% cyber-physical systems.

We have been developing the {\em BriefCASE} toolsuite to address this need.
\brfcs\ is a Model-Based Systems Engineering (MBSE) environment
built in the Open Source AADL\footnote{AADL is the acronym for
Architecture Analysis and Design Language~\cite{aadl}.}  Tool
Environment (OSATE) to add new design, analysis, and code generation
capabilities for building cyber-resilient systems.

In this paper we describe how \brfcs\ facilitates inserting,
specifying, testing, and synthesizing high assurance components into a
system to improve its cyber-resiliency.  The main organizing concept
is that of an architecture-to-architecture \emph{security-improving
transform}, achieved via the insertion of a new architectural
component aimed at mitigating a cyber-vulnerability.  We describe two
cyber-resiliency transformations in this paper: (1) the insertion of a
filter to prevent malformed data from a malicious actor being
propagated to downstream components, and (2) the insertion of a
runtime monitor to detect (and alert) unexpected behaviors arising
from untrusted components.

@@

Another novel aspect of our workthe approach discussed in this manuscript is
the synthesis of the code contracts for the high-assurance components
to \ckml, a verified compiler implementation for the functional
programming language ML \cite{cakeml}.  This manuscript describes in
detail the synthesis path from code contracts to \ckml\ code,
providing a formal framework in which to argue correctness. \ckml\
then provides a verified compilation path to several different target
binaries (and also proving that the meaning of the \ckml\ source code
is exactly preserved in the final binaries).  A key contribution here
is that the code contract semantics are defined in such a way that
the \agr\ verification results for the code contract hold for the
deployed component, \ie, the component will detect and prevent the
indicated cyber-vulnerabilities over all possible finite
inputs. Preliminary work has shown how to lift this result to infinite
input traces as these systems are inherently reactive and intended to
run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.
